Incremental Construction of Programs in a Multitask Setting

### 2013-10-23 drafts/nips/nips2013.tex
Added a description of the combinatory logic.

\documentclass{article} % For LaTeX2e
\usepackage{nips13submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{listings}
%\documentstyle[nips13submit_09,times,art10]{article} % For LaTeX 2.09

\title{Learning Programs through Bayesian Inference and Planning}

@@

Here, we present an approach to learning

Motivation: Many approaches to constructive machine learning use structured representations, such as graphs, grammars, Horn clauses, relations, ..., ???. All of these may be thought of as a program - programs are at least as expressive as these other representations.

% TODO: Sequential EC? SEC? What is a good name for this?
The space of programs is enormous and inherently unpredictable \cite{PACunpredictability}.
We introduce a new algorithm, SEC, for searching the space of programs.
Two main ideas underly SEC:
\begin{itemize}
\item \emph{Learning to search}: Programs admit compositionality and reuse of subprograms. SEC exploits this compositionality to learn a library of program fragments, which it then reuses when performing further searches for programs.
In a multitask setting, this permits bootstrap learning of more complicated programs after the synthesis of simpler programs.
\item \emph{Incremental Program Synthesis}: In many search problems, there is a natural means of assigning partial credit for partially-correct solutions. SEC uses this partial credit to select promising programs and modify them, creating complicated programs step-by-step.
Rather than using a mutation operator, as in genetic programming, or Metropolis-Hastings proposals, as in \cite{MHprogramInduction}, SEC uses function composition to combine smaller programs in to larger ones.
\end{itemize}

The SEC algorithm comes from a generative model over compositions of functions drawn from a stochastic grammar over programs.
We describe this model, as well as how we perform inference, in ...

SEC treats planning as an inference problem.

\section{The SEC Algorithm}
\subsection{Generative Model}
\subsection{Inference}
\subsection{Program Representation}

@@

\begin{figure}
\begin{minipage}[c][8cm][t]{.4\textwidth}
  \vspace*{\fill}
  \includegraphics[width=\linewidth]{./figures/tasks.pdf}
  \caption{test figure one}
  \label{fig:test1}
\end{minipage}%
\begin{minipage}[c][8cm][t]{.3\textwidth}
  \vspace*{\fill}
  \fbox{\includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}}
  \caption{test figure two}
  \label{fig:test2}\par\vfill
  % \includegraphics[width=5cm,height=4.5cm]{im
  \caption{test figure three}
  \label{fig:test3}
\end{minipage}
\end{figure}
\section{Building Towers in Blocks World}

\subsection{The Blocks World Domain}

\subsection{Learned Tower programs}

>>>>

\documentclass{article} % For LaTeX2e
\usepackage{nips13submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{clrscode4e}
\usepackage{tikz}
\usepackage{amsmath}

%\documentstyle[nips13submit_09,times,art10]{article} % For LaTeX 2.09

\title{Incremental Construction of Programs in a Multitask Setting}

@@

\section{The SEC Algorithm}
\subsection{Generative Model}
We propose a generative model over programs that captures two intuitions:
\begin{itemize}
\item{Useful programs are composed of useful subparts.}
\item{Short programs are a prior more likely.}
\end{itemize}
\begin{codebox}
\li \id{G} $\sim$ $P_G(\cdot)$ \Comment{Draw grammar from description-length prior}
\li \id{\ell_n} $\sim\mbox{Uniform}(1, L)$ \Comment{Draw number of subprograms in $n^{th}$ program}
\li \Comment{Draw $\rho_n$, the $n^{th}$ program, by composing its subprograms, $e_n^{i}$}
\li \id{e_n^{i}} $\sim P_{e|G}(\cdot | G)$
\li \id{\rho_n} $= e_n^{\ell_n} (e_n^{\ell_n-1} ( \cdots e_n^{1}))$
\li \Comment{Draw $t_n$ by adding noise $\epsilon_n$ to the output of $\rho_n$}
\li \id{\epsilon_n} $\sim P_{\mbox{noise}}(\cdot)$
\li \id{t_n} $= \rho_n + \epsilon_n$
\end{codebox}

Write down the generative model. Connect to EC and Percy Liang's work. Discuss ``programs as plans.''
\subsection{Inference}
We implement an iterative Expectation-Maximization algorithm that performs MAP estimation of the libraray, $G$, given the tasks, $\{t_n\}$.
The joint distribution factorizes as
\begin{equation}
P(G,\{\rho_n\},\{t_n\}) = \frac{1}{L^N} P(G) \prod_n P(t_n | \rho_n) \prod_i P(e^i_n | G)
\label{eq:joint}
\end{equation}
From equation (\ref{eq:joint}), the EM updates are
\begin{eqnarray}
q(\rho_n) &\propto& P(t_n | \rho_n) \prod_i P(e^i_n | G^{old})\\
\label{eq:qdist}
G^{new} &=& \operatorname{argmin}_G \left( -\ln P(G) -
\sum_n 
E_{q_n}
 \left[ \ln P(\rho_n | G) \right] \right)
 \label{eq:gmax}
\end{eqnarray}
Exact computation of $q(\cdot)$ in equation (\ref{eq:qdist}) would involve marginalizing over the space of all programs.
Instead, we perform a heuristic search over the space of programs, with the goal of finding those programs for which $q(\cdot)$ is high.
Rather than use mutation and crossover operators, as done in genetic programming, or random resampling of subprograms, as done in recent work in applying Metropolis-Hastings to program learning, we propose that the search moves should themselves be programs drawn from the grammar $G$.
Concretely, in order to modify a program $e$ to produce new candidate programs, we draw a new function from our grammar, $G$, and apply that new function to $e$.
Thus, we use the grammar not only as a distribution over programs of interest, but also as a distribution over programs that take as input other programs.
This incremental construction of programs by repeatedly applying draws from the grammar mirrors the construction of programs in the generative model.

In our experiments, we used beam search. The objective function was the unnormalized posterior $P_{t_n|\rho_n}(t_n | \cdot )P_{\rho_n | G}(\cdot | G)$.
The beam was initialized to be the programs with the highest prior probability according to the grammar $G$.
Similarly, the search moves considered were those functions with the most highest prior probability under $G$.

Equation (\ref{eq:gmax}) has a natural interpretation as a form of compression.
Interpreting negative log probabilities as description lengths, the update (\ref{eq:gmax}) picks the grammar minimizing the sum of the description length of the grammar, plus the expected description length of the programs found to solve the tasks.

Performing the minimization operation in equation (\ref{eq:gmax}) is generally difficult.
We approximate the description length of the grammar by the number of unique program fragments in the grammar, multiplied by a regularization constant.
The description length of a program is approximated by the number of (not necessarily unique) program fragments in its smallest parse tree given the grammar.
A slight generalization of the Neville-Manning algorithm then permits tractible minimization of (\ref{eq:gmax}).
Estimation of the production probabilities of $G$ is performed as in \cite{ijcai}.

\subsection{Program Representation}

Following~\cite{DBLP:conf/ijcai/DechterMAT13} and~\cite{DBLP:conf/icml/LiangJK10}, we use a
simply typed combinatory calculus -- a variable-free subset of the
polymorphic simply-typed lambda calculus~\cite{DBLP:books/daglib/0005958} -- as our
program representation language. The simplicity the combinatory calculus makes describing a distribution over its expressions particularly simple. A more detailed explanation and discussion of this program representation can be found in the papers cited above.

\section{Application to Structure Learning}
Learning the structure of graphical models is a difficult problem - not only for machines, but also for humans, where expert knowledge is often required.
In practice, the sorts of graphical models that humans design, such as Hidden Markov Models, phylogenetic trees, topic models, and Ising models, exhibit certain symmetries and recursive structures that are amenable to synthesis by programs.
Could an algorithm, such as SEC, automatically learn these structures?

@@

Show the actual results for our test cases: ``density estimation of the space of interesting graphical models.''
In future work, could be used as high-level moves in the search for a good structure.
Connect to, eg, Charles Kemp's work.

\begin{figure}
\begin{minipage}[c][8cm][t]{.4\textwidth}
  \vspace*{\fill}
  \includegraphics[width=\linewidth]{./figures/tasks.pdf}
  \caption{test figure one}
  \label{fig:test1}
\end{minipage}%
\begin{minipage}[c][8cm][t]{.3\textwidth}
  \vspace*{\fill}
  \fbox{\includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}}
  \caption{test figure two}
  \label{fig:test2}\par\vfill
  % \includegraphics[width=5cm,height=4.5cm]{im
  \caption{test figure three}
  \label{fig:test3}
\end{minipage}
\end{figure}
